### 2012-03-24 comparison.tex
breaking into separate files

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{natbib}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amsmath}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Graphical Tests for Power Comparison of Competing Designs}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Lendie Follett, Heike Hofmann, Mahbub Majumder, Dianne Cook}
\authorfooter{
%% insert punctuation at end of each item
\item
 Lendie Follett is a graduate student in Statistics at Iowa State University, E-mail: lfollett@iastate.edu.
\item
everybody else is at Iowa State University, too
}

%other entries to be set up for journal
\shortauthortitle{Follett \MakeLowercase{\textit{et al.}}: Comparison of Competing Layouts}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{
Lineups  \cite{buja:2009, wickham:2010} have been established as tools for visual testing
similar to standard statistical inference tests, allowing us to
evaluate the validity of graphical findings in an objective manner. In
simulation studies  \cite{majumder:2011} lineups have been shown as being efficient: the
power of visual tests is comparable to classical tests while being
much less stringent in terms of distributional assumptions made. This
makes lineups versatile, yet powerful, tools in situations where
conditions for regular statistical tests are not or cannot be met. In
this paper we introduce lineups as a tool for evaluating the power of
competing graphical designs. We highlight some of the theoretical
properties and then show results from two studies evaluating competing
designs: both studies are designed to go to the threshold of our
perceptual abilities to highlight differences between designs. We use
both accuracy and speed of evaluation as measures of a successful
design. The first study compares the choice of coordinate system:
polar versus Euclidean coordinates. The results show strong support in
favor of Euclidean coordinates in finding fast and accurate answers to
spotting patterns. The second study is aimed at finding shift
differences between distributions. Both studies are motivated by data
problems that we have recently encountered, and explore using
simulated data to evaluate the plot designs under controlled
conditions.  Amazon Turk is used to conduct the studies.  The lineups
provide an effective mechanism for objectively evaluating plot
designs.
} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Lineups, visual inference, Power Comparison, Efficiency of Displays.}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{ % not used in journal version
  \category{H.5.2}{Information Interfaces and Presentation}{ User Interfaces --- Graphical user interfaces (GUI), Interaction styles, Screen design, Evaluation/methodology}
  \CCScat{I.6.8}{Computing Methodologies}%
  {Simulation and Modeling}{Visual Simulation};
}

\graphicspath{{images/}}

%% Uncomment below to include a teaser figure.
\teaser{
\centering
\includegraphics[height=1in]{polar}\hspace{0.05\linewidth}
\includegraphics[height=1in]{eucline}
%\includegraphics[width=16cm]{CypressView.eps}
\caption{teaser chart should include all different designs}
}

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

\input{intro}

\input{experiment}

\section{Conclusion}

%% if specified like this the section will be ommitted in review mode
\acknowledgments{
This work was supported in part by
NSF DMS 1007697.}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{references}
\end{document}