### 2012-03-29 conclusions.tex
rearranged paper

\section{Conclusions}~\label{conclusions}

Could say something about adding a visual pleasure rating to the results, rate how attractive the display is on a scale of 1-5.