### 2012-03-30 conclusions.tex
some more on confidence levels in experiment 2

\section{Conclusions}~\label{conclusions}

Could say something about adding a visual pleasure rating to the results, rate how attractive the display is on a scale of 1-5.

Familiarity with the design?

Assessing confidence leads to interesting discrepancies between perceived versus actual accuracy. In both examples, the only significant increases in confidence were not indicative of an increase in actual performance, in one case even the opposite happened.