### 2012-06-27 review/response.txt
revision

I would like to thank the reviewers for the great discussion of the paper, and a lot of helpful suggestions!

We have made the following changes:

All of the main suggestions are implemented:
 - in the presentation of figure 8, differences between power results are now presented using a Benjamini-Hochberg adjustment for the false discovery rate.
   Since BH adjustments are based on individual p-values (whereas the Bonferroni adjustment results in a global scaling of the intervals), it was necessary to change the presentation in the figure slightly - we now show the significant differences using a letter-based approach as suggested by H.-P. Piepho (2004).

 - the paper is re-organized and the two studies are now presented one after the other.

 - the discussion regarding pie vs barcharts is removed - we'd like to thank the reviewers for pointing out the additional references.

Changes based on suggestions of individual reviewers:
 - the blue scale for the pictures of the wind direction was darkened a bit, which makes the small levels better distinguishable both in print and pdf.

 - reporting the log: we indeed used log base 10 for figure 6 - we changed this picture to natural log for consistency.

 - Euclidean vs Cartesian coordinates - agreed, and changed

 - language concerning null hypothesis and alternative in 3.2 is fixed.

 - The scale for the self-reported confidence is included explicitly now.

 -  the design choices for the second study were based on designs commonly used; we have addressed this in the paper.

 - conclusion section is strengthened and also focuses on the results from the experiment rather than just their setup.

Discussion:

Q: Would it be more understandable and specific to characterize the power
  model as a hierarchical linear model with random intercept instead of a
  generalized linear model? If I understand correctly, the HLM with random
  intercept is specifically what they are using. If not, the authors can
  explain to the reviewers why I am wrong and not make the change.

A: What we are using in estimating power is actually a generalized linear model with a random intercept - since power has to be a value between 0 and 1, we are using a logistic regression instead of a linear model.

Q: The first study does not state clearly what constitutes "best" time for
  minutes between wheel events. You claim that it is clear that Seattle
  functions at its best with winds coming from the East but I'm not sure if
  that means that you just want the lowest number of minutes or if there is
  an optimal number. The hypotheses just address "differences" but not any
  specific time interval in general.

A: We were trying to avoid the phrase 'best' - as we are, indeed, just looking at efficiency ('its best' has been changed in the paper to 'most efficiently') - i.e. less time between wheel events is considered 'better'. (Note: Seattle has several runways, so a time of 0 between wheel events is not as unsafe as it might sound).

Hans-Peter Piepho (2004), An Algorithm for a Letter-Based Representation of All-Pairwise Comparisons, Journal of Computational and Graphical Statistics, 13(2), 456–466.