### 2012-06-27 review/experiment-two.tex

% !TEX root = comparison-review.tex
\subsection{Study II: Comparing Distribution Centers}
\subsubsection{Study Setup}
In a second experiment, we conducted a simulation study to investigate the power of four competing designs, as shown in Figure \ref{fig:expii}, in assessing a mean shift between distributions.

\noindent The  hypotheses for each lineup are

\begin{tabular}{rl}
$H_o:$&{\em centers of the two groups are the same}, \\
$H_a:$&{\em center of the blue group is shifted to the right}.\\[3pt]
\end{tabular}
%\vspace{-0.1in}
%\vspace{-0.1in}

As covariates for this experiment, we considered the {\it size of the shift}, $d$, between the means of the distributions, with $d=0.4, 0.6, 0.8. 1.0,$ and $1.2$, as well as the sample size.  Since we are dealing with two groups of points, we varied both the {\em size of the larger group}, $n_1 = 15, 45$, and 135, as well as the {\em relative size} of the second group ({\em ratio}), with $n_2 = 1/3, 2/3$ and $3/3$ of $n_1$. 
For each of these combinations, we created three replicate data sets with values from an exponential distribution with rate parameter $\lambda=1$ (for the first group) and rate parameter $\lambda_2=1/(d+1)$, which corresponds to a mean shift of $d$, for the second groups. 
Based on each data set we created four lineup charts, one for each of the competing designs. This resulted in 540 different lineup charts overall, which we sorted by $p$-value corresponding to the difference in means between the two groups in the simulated data. According to the $p$-value, we associated a `difficulty' level with lineups from 1 to 9 (easy to hard). Ten lineup plots were shown to each participant - one out of each of the difficulty levels and one additional reference chart with a particularly easy lineup  to allow for data quality checks.
\begin{figure} [hbtp]
   \centering
   \includegraphics[width=0.45\linewidth]{hist-conc} 
   \includegraphics[width=0.45 \linewidth]{boxplot-conc} \\
   \includegraphics[width=0.45 \linewidth]{density-conc} 
   \includegraphics[width=0.45 \linewidth]{dotplot-conc} 
\vspace{-0.1in}
   \caption{Overview of all four competing designs for displaying differences in distributions.}
   \label{fig:expii}
\end{figure}

All of the designs were chosen based on common usage. The main difference between the designs is their varying level of data summarization.

 From most to least summarized we consider these designs:

\begin{tabular}{lp{0.33\textwidth}}\\[-5pt]
\em boxplots: & five-point summary statistics (and a number of  outliers),\\
\em density plots: & need -- mathematically -- all of the data points, but  their smoothness and continuity makes them appear visually quite simple,\\
\end{tabular}
\begin{tabular}{lp{0.33\textwidth}}
 {\em histograms}: &  rendered by a frequency breakdown for a specified number of bins ($10 \times 2$, in the study),\\
 {\em dotplots}: & all $n$ data points necessary for rendering.\\[3pt]
\end{tabular}

\noindent For the boxplot and dotplot, the vertical axis is free to be used for another variable, {\em group}, so the data can be split apart. For the histogram and density the vertical axis is used for count or frequency, and is not available for group information.
Based on the assumption that charts are most effective \citep{Kosslyn:2006}, when they show enough information to solve the task at hand, but not too much to hamper the observer, we anticipated results to reflect the above order.

\subsubsection{Results}

Since each participant provides results from nine lineups, we can use a generalized linear mixed effects model of power and account for individuals' abilities by including a subject-specific random intercept. All of the above described covariates are considered  in the model, both as main effects and as two-way interaction with the design to assess their impact on the power of each of the designs.
Figure \ref{fig:power2} summarizes the modeling results as shown in table \ref{tbl:power2}. Generally we can see that the power of all four designs increases with an increase in effect size $d$ (on the $x$ axis), as well as an increasing sample size $n_1$ (top to bottom facetting). Generally, an increase of the second group also is beneficial to power. Dotplots are an exception to this -- they actually lose power significantly as the second group increases. This might be due to visual complexity, as discussed above.
%The reason behind this phenomenon might be visual complexity \citep{Kosslyn:2006}: dotplots are the only design that show individuals. With an increase in group size, visual complexity increases with it, which could hamper the observer in evaluating shifts.

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\begin{tabular}{lrrrrrl}
  \hline
& & Estimate & Error & $t$-value & \multicolumn{2}{l}{approx $p$-value} \\   \hline
\\[-8pt]
\bf design& Intercept & -3.11 & 0.41 & -7.61 & 0.00 & ***\\ 
&  density & 0.05 & 0.57 & 0.09 & 0.93 \\ 
&  dotplot & 0.46 & 0.57 & 0.81 & 0.42\\ 
&  histogram & 0.06 & 0.58 & 0.11 & 0.91 \\ [2pt]
\multicolumn{2}{l}{\bf covariates}\\
&  d & 1.77 & 0.34 & 5.18 & 0.00 & ***\\ 
&  n1$_{100}$ & 1.84 & 0.19 & 9.73 & 0.00 & ***\\ 
&  ratio & 0.84 & 0.33 & 2.56 & 0.01 & **\\ 
\multicolumn{2}{l}{\bf interactions}\\
&  density:d & -0.60 & 0.47 & -1.27 & 0.20 \\ 
&  dotplot:d & 0.03 & 0.46 & 0.07 & 0.95  \\ 
&  histogram:d & -0.28 & 0.48 & -0.59 & 0.56 \\ [1pt]
&  density:n1$_{100}$ & -0.10 & 0.26 & -0.36 & 0.72\\ 
&  dotplot:n1$_{100}$ & -0.77 & 0.26 & -3.00 & 0.00 & ***\\ 
&  histogram:n1$_{100}$ & -0.69 & 0.26 & -2.64 & 0.01 & ** \\ [1pt]
&  density:ratio & -0.28 & 0.47 & -0.59 & 0.56 \\ 
&  dotplot:ratio & -1.01 & 0.46 & -2.19 & 0.03 & * \\ 
&  histogram:ratio & -0.28 & 0.47 & -0.59 & 0.55 \\ 
   \hline
\\[-5pt]
   \multicolumn{5}{l}{Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1}
\end{tabular}}
\end{center}
\vspace{-0.2in}
\caption{\label{tbl:power2}Results from a Generalized Linear Mixed Effects Model of power of a  lineup given design and all two-way interactions with effect size $d$, group size $n_1$, and ratio to second to first group size. 
Note that for all terms involving $n_1$ estimate and error are multiplied by a factor of 100. 
Reference group for all effects are boxplot designs. The model fit is based on 2513 lineup evaluations by 208 participants. }
\end{table}

Visually, density plots seem to 
The two highly summarized designs benefit the most from an increase in sample size: both density charts and boxplots profit significantly more than histograms and dotplots. This might, again, be a hint towards a trade-off between visual complexity and amount of information shown as the basis for making decisions.

Table \ref{tbl:speed2} gives an overview of results from a linear model for (log) evaluation time of lineups. Subject-specific speed (due to spatial ability) are accounted for  including a random intercept. Structurally, the model is the same as before, i.e. we regard effect size $d$, size $n_1$ of the first group, and the ratio $n_2/n_1$ between group sizes both as main effects and in two-way interactions with the four competing designs. To make estimate sizes comparable, we have again multiplied all estimates and standard errors involving the group size $n_1$ by 100. This means that we need to interpret those  estimates as average change if the group size is increased by 100 elements.

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\begin{tabular}{lrrrrrl}
  \hline
& & Estimate & Error & $t$-value & \multicolumn{2}{l}{approx $p$-value} \\   \hline
\\[-8pt]
\bf design& Intercept & 3.85 & 0.09 & 42.21 & 0.00 & ***\\ 
&  density & 0.01 & 0.12 & 0.07 & 0.94 \\ 
&  dotplot & 0.03 & 0.12 & 0.26 & 0.80 \\ 
&  histogram & 0.23 & 0.12 & 1.97 & 0.05 & . \\ [2pt]
\multicolumn{2}{l}{\bf covariates}\\
&      d & -0.23 & 0.08 & -3.02 & 0.00 & ***\\ [1pt]
&    n1$_{100}$ & 0.07 & 0.04 & 1.70 & 0.09 & .\\ [1pt]
&  ratio & -0.04 & 0.02 & -1.48 & 0.14  \\ [2pt]
\multicolumn{2}{l}{\bf interactions}\\
&  density:d & 0.11 & 0.11 & 1.02 & 0.31 \\ 
&  dotplot:d & 0.20 & 0.11 & 1.86 & 0.06 & .  \\ 
&  histogram:d & -0.01 & 0.11 & -0.09 & 0.93 \\ [1pt]
&  density:n1$_{100}$ & -0.12 & 0.06 & -2.11 & 0.03 & *\\ 
&  dotplot:n1$_{100}$ & -0.16 & 0.06 & -2.83 & 0.00 & ***\\ 
&  histogram:n1$_{100}$ & -0.05 & 0.06 & -0.93 & 0.35 \\ [1pt]
&  density:ratio & 0.04 & 0.03 & 1.18 & 0.24 \\ 
&  dotplot:ratio & 0.02 & 0.03 & 0.65 & 0.52 \\ 
&  histogram:ratio & 0.00 & 0.04 & 0.04 & 0.97 \\ 
   \hline
\\[-5pt]
   \multicolumn{5}{l}{Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1}
\end{tabular}}
\end{center}
\vspace{-0.2in}
\caption{\label{tbl:speed2}Results from a Linear Mixed Effects Model of (log) time to answer a lineup given design and all two-way interactions with effect size $d$, group size $n_1$, and ratio to second to first group size. Estimate and error for all terms involving $n_1$ are multiplied by a factor of 100 for comparability.
Reference group for all effects are boxplot designs. }
\end{table}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.95\linewidth]{conf-exp2} 
\vspace{-0.25in}
   \caption{Barcharts of marginal distributions of self-reported confidence by design. The distributions display prominent differences.}
   \label{fig:conf-margins}
\end{figure}

The results of this model are similar to the previous: evaluation speeds needed for the four different designs is very close, but  histograms take on average a significantly longer time ($e^{0.23} = 1.26$ seconds) to evaluate than lineups of boxplots. 
Where an increase of sample size led to an increase of accuracy before, we now see that it increases time for evaluation of boxplot slightly, but not significantly so. What is interesting is that evaluation time decreases significantly with an increase in sample size, particularly for dotplots, but also for density charts. This is counter-intuitive to our previous argument on dotplots being affected by visual complexity the most. It seems, that observers of dotplot lineups come to a conclusion faster as sample size increases, albeit more often to  the wrong conclusion. Rather than visual complexity, this might be indicative that over-plotting is problematic in dotplots. While we tried to accommodate for over-plotting by making use of alpha-blending and jittering, this might not have been sufficient for the large sample size.

Our third response variable is again the subjective assessment of confidence level. On average participants reported a confidence in their results of  3.41 (on a five point scale with 1=lowest and 5=highest), with a standard deviation of 0.149. Modeling confidence levels with a linear mixed effects model with the same structure as before resulted in only one significant effect beyond the intercept: participants report significantly higher confidence for increased number of  points in dotplot lineups. On average an increase of 100 points leads to an increase in confidence of 0.22 (with a standard error of 0.082). This finding is significant with a $p$-value of 0.007. 
Together with the previous results, we can tell that something worthy of further investigation happens in the perception of dotplots as sample sizes increase.

Figure \ref{fig:conf-margins} shows barcharts of confidence assessments by design. It is obvious that the distribution of confidence is very different from one design to the next. Boxplots and dotplots show a similar pattern with an emphasis on high confidence levels, whereas histograms and density lineups have modes at a confidence level of three. 
One reason behind these differences in confidence distributions might be familiarity with the design - neither density charts nor histograms are particularly frequent displays for a general audience.